---
title: "GGPlot2 Chapter 12"
author: "Hannah Gelman"
date: '2020-05-04'
output:
  html_document:
    df_print: paged
---

### Notebook with Notes and Exercises from Chapter 12.  

```{r}
library(ggplot2) 
library(tidyverse)
```
```{r}
head(mpg)
```

##### 12.1 Specifying scales 

Every aesthetic in your plot is mapped to a single scale. 

Only one scale is allowed/aesthetic, so subsequent scale setting on the same aesthetic effectively overrides the previous setting. 

1. What happens if you pair a discrete variable to a continuous scale or vice-versa? 
--> error. discrete value supplied to continuous scale or 
--> Error: Continuous value supplied to discrete scale --> this seems to be a little inconsistent. in all cases it is difficult for labels to be associated with the scale. sometimes points seem to be plotted but in others an error is thrown

```{r}
ggplot(mpg, aes(year, displ)) + geom_point() + scale_x_discrete()
```

```{r}
ggplot(mpg, aes((hwy), displ)) + 
  geom_point() + 
  scale_x_discrete(limits = as.character(c(12:44)))

```

2. Simplify the following plot specifications 
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() +
  scale_y_continuous("Highway mpg") 
```
```{r}
ggplot(mpg, aes(class, displ, color = drv)) + 
  geom_point() + 
  scale_y_continuous("Displacement (l)") + 
  scale_x_discrete("Type of car") + 
  scale_colour_discrete("Drive\ntrain")
```

##### 12.2 Adjusting scale limits 

ggplot scales are usually derived from data min and max. you can specify specific limits to: 
1. shrink to focus on an interesting area of the plot 
2. expand to make multiple plots match up. 

To change these, you use the limits argument in the scale_ functions. for a continuous function the limits are a numeric vector  = c(min, max), for discrete scales, limits is a character vector enumerating all possible values. 

xlim, ylim exist as convenience functions for this. lims is a general function 

to eliminate expansion of limits, use expand  = c(0,0) 

Exercises 

```{r}
fwd <- subset(mpg, drv == "f")
rwd <- subset(mpg, drv == "r")

# set up limits 
hwy <- c(unique(rwd$hwy), unique(fwd$hwy)) 
class <- unique(c(rwd$class, fwd$class))
displ <- unique(c(rwd$displ, fwd$displ))


```

```{r}
ggplot(fwd, aes(displ, hwy, colour = class)) + 
  geom_point() +
  scale_color_discrete(limits = class) +
  lims(x = c(min(displ), max(displ)), 
       y = c(min(hwy), max(hwy)))
```

```{r}
ggplot(rwd, aes(displ, hwy, colour = class)) + 
  geom_point() +
  lims(x = c(min(displ), max(displ)), 
       y = c(min(hwy), max(hwy)), 
       color = class)
```

