---
title: "GGPlot2 Chapter 13"
author: "Hannah Gelman"
date: '2020-06-14'
output:
  html_document:
    df_print: paged
---

####Chapter 13: Guides: legends and axes 

Legends and axes both map to data and help you read the original values off the plot. 

In ggplot2, you have relatively little control of the legend, and should set up your data to get the legend you want. 

legends and axes are controlled by the same set of arguments: 
1. name (title) 
2. breaks (ticks on axis, key in legend)
3. labels (tick label on axis, key labe in legend. )

Though share features, legend is more complicated than axis because it can display multiple aesthetics applied across multiple layers. Also, more options for formatting and display.  

```{r}
library(ggplot2)
library(tidyverse)
```


#####13.1-13.2 Basic Scale Controls 

The first argument in the scale functions is "name". Special functions xlab, ylab, labs

`breaks` controls which values appear. `labels` controls what is displayed. 

if you manually set `labels`, you must also set `breaks` otherwise the twho will be out of sync. 
#####13.2.2 Execises 

```{r}
head(mpg)
```


1. Recreate the graphic 
```{r}
mpg %>% ggplot(aes(displ, hwy)) + 
  geom_point(size = 3) + 
  scale_x_continuous("Displacement", 
                     labels = scales::unit_format(accuracy = .1, 
                     suffix = "L")) + 
  scale_y_continuous(expression(paste("Highway ", bgroup("(", frac(miles, gallon), ")"))))
```

2. List the 3 different types of objects you can supply to the breaks argument
- a vector 
- a function that generates breaks from limits (suggested a 2 element of limits?)
- NULL (to suppress) (?) 

`labels` argument does not need to map to data, but describes how the breaks will be labeled. 

3. Recreate the following plot 
```{r}
mpg %>% ggplot(aes(displ, hwy, color = drv)) + 
  geom_point(size = 3) + 
  scale_color_discrete(labels = c('4wd', 'fwd', 'rwd'))
```

4. What label function allows you to create mathematical expressions? `expression` 
What label function allows you to convert to ordinal numbers `scales::label_ordinal`

5. `name`, `breaks`, `labels`. I'm not explaining, that's just rewriting what's above. 

#####13.2.3 Dates

There are special helper functions for formatting dates. remember them, i guess? 

#####13.5.2 Legends Exercises 

1. How do you make legends appear to the left of the plot?
`theme(legend.position = "left")`

```{r}
df <- data.frame(x = 1:3, y = 1:3, z = c("a", "b", "c"))
base <- ggplot(df, aes(x, y)) + 
  geom_point(aes(colour = z), size = 3) + 
  xlab(NULL) + 
  ylab(NULL)

base + theme(legend.position = "left")
```

2. What's gone wrong with this plot? How could you fix it? 
2 aesthetics matched to drv getting disconnected because name of only one scale (color) has been changed. changed the name of shape scale as well to re-link

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = drv, shape = drv)) + 
  scale_colour_discrete("Drive train") + 
  scale_shape_discrete("Drive train")
```

3. Can you recreate the code for this plot
```{r}
ggplot(mpg, aes(displ, hwy, color = class)) + 
  geom_point(show.legend = FALSE) + 
  geom_smooth(method = lm, se = FALSE) + 
  scale_color_discrete(name = NULL) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 1))

```

